---
title: "Regionalisation of the Environment"
author: "Markus Diesing"
output: 
  html_document: 
    toc: yes
---

# Packages

```{r packages, echo=TRUE, message=FALSE, warning=FALSE}
if(!require("pacman")) install.packages("pacman"); library(pacman)
p_load(here,
       terra,
       sf,
       RStoolbox,
       ggplot2,
       ggpubr,
       gridExtra
       )
```


# Loading data

## Creating a rasterstack

```{r rasterstack, warning=FALSE}
cri <- rast(here("data", "raw", "CRI_median_2024-06-21.tif"))
ocar <- rast(here("data", "raw", "OCAR_2024-03-20.tif"))
ocs <- rast(here("data", "raw", "OCS0-10cm_2024-03-20.tif"))

vars <- c(ocs, ocar, cri)
names(vars) <- c("OCS", "OCAR", "CRI")
```


## shp files 

These will be used at the end to create a figure.

```{r shp}
land<- read_sf(here("data", "raw", "continent.shp"))
land <- st_transform(land[3,], crs(vars))
study_site <- st_transform(read_sf(here("data", "raw", "AoI_Harris_mod_dissolve.shp")), crs(vars))
```



# Normalise raster layers

Performs normalisation (scaling) and centering, i.e. divide by standard deviation

```{r normalisation}
vars.norm <- normImage(vars, norm = TRUE)
vars.norm
hist(vars.norm)
```


# kmeans classification 

https://www.rdocumentation.org/packages/RStoolbox/versions/0.2.3/topics/unsuperclass

## Specify the maximum number of clusters to be analysed

```{r max_number_of_clusters}
max_cl <- 12
```

## How many clusters?

```{r how_many_clusters}
df_total <- data.frame()
for (i in 1:max_cl) {
  set.seed(25)
  unC <- unsuperClass(vars.norm[[1:nlyr(vars.norm)]], nSamples = 4000, nClasses = i, nStarts = 100, 
                      norm = FALSE, clusterMap = TRUE, algorithm = "Hartigan-Wong")
  
  stat <- (unC[["model"]][["tot.withinss"]])
  df <- data.frame(i,stat)
  df_total <- rbind(df_total,df)
}
```

## Plot graph

```{r graph}
elbow_plot <- ggplot(df_total,aes(i,stat)) + geom_bar(stat = "identity") +
              xlab("Number of clusters") + ylab("total within sum of squares ") +
              scale_x_continuous(breaks = c(1:max_cl))
elbow_plot

ggsave(filename = "elbow_plot.jpg", plot = elbow_plot, device = "jpeg", path = here("figures"), height = 10, width = 15, units = "cm", dpi = 500)
```

## Specify the number of clusters to be used

based on the graph above

```{r select_number_of_clusters}
n.cl <- 6
```

## Run k-means clustering

```{r k_means}
set.seed(25)
unC <- unsuperClass(vars.norm, 
                    nSamples = 10000, 
                    nClasses = n.cl, 
                    nStarts = 100, 
                    norm = FALSE, 
                    clusterMap = TRUE, 
                    algorithm = "Hartigan-Wong")

unC$model
```


## Change cluster number by means of OCAR

```{r change_cluster_number}
lut <- matrix(nrow = n.cl, ncol = 2)
lut[,1] <- as.numeric(names(sort(unC$model$centers[,2], decreasing = TRUE)))
lut[,2] <- c(1:n.cl)

regions <- classify(unC$map, lut)
```


## Define colour palette

```{r col_pal}
col_pal <- c('#1f78b4', '#a6cee3','#b2df8a','#33a02c', '#fb9a99', '#e31a1c', '#fdbf6f', '#ff7f00', '#cab2d6', '#6a3d9a', '#ffff99', '#b15928')
```


## Plot regions

```{r plot_regions, warning=FALSE}
ggR(regions, geom_raster = TRUE, forceCat = TRUE) +
  scale_fill_manual(values = col_pal, na.value = NA)
```


# Export Geotiff

```{r export_geotiff, warning=FALSE}
writeRaster(regions, here("data", "ready", paste0("regions_", n.cl, ".tif")), overwrite = TRUE)
```

# Violin plots

```{r violin plots}
r <- c(regions, vars)
smp <- as.data.frame(spatSample(x = r, size = 10000, method = "random", na.rm = TRUE))
smp[,1] <- as.factor(smp[,1])
names(smp)[1] <- "Cluster"

for (i in 2:ncol(smp)) {
  print(ggplot(smp, aes(x = Cluster, y = smp[,i], fill = Cluster)) +
  geom_violin() +
 scale_fill_manual(values = col_pal) +
 scale_y_continuous(name = names(smp[i])) +
 theme(axis.title.x = element_blank()) +
  theme(legend.position = "none"))
  }
```

# Figure for publication

```{r figure, warning=FALSE}
AoI <- ext(regions)
land <- st_crop(land, AoI)

p0 <- ggR(regions, geom_raster = TRUE, forceCat = TRUE) + 
  geom_sf(data = study_site, fill = NA, linetype = 5) +
  geom_sf(data = land, fill = "darkgrey") +
  scale_fill_manual(values = col_pal, na.value = NA) +
  theme_bw() +
  theme(axis.title = element_blank()) +
  theme(legend.position = "none" ) + 
  ggpubr::rotate_y_text()

p1 <- ggplot(smp, aes(x = Cluster, y = CRI, fill = Cluster)) +
  geom_boxplot() +
  scale_fill_manual(values = col_pal) +
  scale_y_continuous(name = "CRI (-)") +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none")
p2 <- ggplot(smp, aes(x = Cluster, y = OCAR, fill = Cluster)) +
  geom_boxplot()+
  scale_fill_manual(values = col_pal) +
  scale_y_continuous(name = expression(OCAR~(g~m^{-2}~yr^{-1}))) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none")
p3 <- ggplot(smp, aes(x = Cluster, y = OCS, fill = Cluster)) +
  geom_boxplot()+
  scale_fill_manual(values = col_pal) +
  scale_y_continuous(name = expression(OC~stock~(kg~m^{-2}))) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none")

lay <- rbind(c(1,3),
             c(1,2),
             c(1,4))

p <- grid.arrange(p0, p3, p2, p1, nrow=4, layout_matrix=lay)
ggsave( paste0("regions_", n.cl, ".jpg"), plot = p, device = "jpeg", path = here("figures"), height = 20, width = 20, units = "cm", dpi = 500, type ="cairo")
```
